---
title: "All Cells Recursive Analysis"
author: "Paul Hook"
output: html_document
---

**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

###Setting important directories. Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))
```

###Removing outliers found in recursive analyses
```{r Remove outliers}
# Removing outliers from other analyses
P7.outliers <- readRDS(file = file.path(Rdata_dir,"P7.outliers.rds"))

dat.filter <- readRDS(file = file.path(Rdata_dir, "dat.filter.rds"))
nrow(pData(dat.filter)) # 410

dat.filter <- dat.filter[,!(pData(dat.filter)$sample_id.x %in% P7.outliers)]
nrow(pData(dat.filter)) # 396
```

###Filter genes by percentage of cells expresssing each gene
```{r filter by cells expressing each gene}
# Plot number of cells expressing each gene as histogram
hist(fData(dat.filter)$num_cells_expressed,
     breaks=50,col="red",main="Cells expressed per gene")

# Keep only expressed genes with expression in >= 5% of cells
numCellThreshold<-nrow(pData(dat.filter))*0.05
dat.expressed_genes<-row.names(subset(fData(dat.filter),num_cells_expressed >= 5))

# Same plot as above with threshold
hist(fData(dat.filter)$num_cells_expressed,breaks=50,col="red",main="Cells expressed per gene - threshold")
abline(v=5,lty="dashed")
```

###Prepping the Monocole model for analysis
```{r model_prep}
# Preparing the model for monocle
dat.filter <-dat.filter[dat.expressed_genes,]

dat.filter <-estimateSizeFactors(dat.filter)

dat.filter <- estimateDispersions(dat.filter,cores=8)
# Removing 144 outliers
# Warning message:
# Deprecated, use tibble::rownames_to_column() instead. 
```

###Calculating summary stats
```{r summary_stats}
# Calculating summary stats
fData(dat.filter)$mean_expr<-apply(round(exprs(dat.filter)),1,mean)
fData(dat.filter)$sd_expr<-apply(round(exprs(dat.filter)),1,sd)
fData(dat.filter)$bcv<-(fData(dat.filter)$sd_expr/fData(dat.filter)$mean_expr)**2
fData(dat.filter)$percent_detection<-(fData(dat.filter)$num_cells_expressed/dim(dat.filter)[2])*100
```

###Identifying high dispersion genes
```{r high_dispersion_genes_monocle}
dat.filter.genes <- dat.filter
disp_table <- dispersionTable(dat.filter.genes)
unsup_clustering_genes <-subset(disp_table, mean_expression >= 0.5 & dispersion_empirical >= 1.5 * dispersion_fit)
dat.high_bcv_genes<-unsup_clustering_genes$gene_id
dat.filter.order <- setOrderingFilter(dat.filter, unsup_clustering_genes$gene_id)
plot_ordering_genes(dat.filter.order)
length(dat.high_bcv_genes) # 1071
```

###Running PCA with high dispersion genes
```{r Run PCA with high BCV genes}
# BCV Identified Marker genes
dat.filter.BCV.pca<-prcomp(t(log2(exprs(dat.filter[dat.high_bcv_genes,])+1)),center=T,scale. = TRUE)
saveRDS(object = dat.filter.BCV.pca, file.path(Rdata_dir, "dat.filter.BCV.pca"))

# Plot the PCA plots
hvPCA1<-ggbiplot(dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(dat.filter)$age,ellipse=T,var.axes=F) + scale_color_manual(values=c("darkgreen","red")) + monocle:::monocle_theme_opts() 

hvPCA2<-ggbiplot(dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(dat.filter)$region,ellipse=T,var.axes=F) + scale_color_brewer(palette="Set1") + monocle:::monocle_theme_opts()

hvPCA3<-ggbiplot(dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(dat.filter)$split_plate,ellipse=T,var.axes=F) + scale_color_brewer(palette="Set1") + monocle:::monocle_theme_opts()

# Show the plots
hvPCA1
hvPCA2
hvPCA3
```

###Viewing screeplots and determining the number of "significant" PCs.
```{r Screeplots of Data,tidy=TRUE}
# Making a screeplot of the BCV PCA. This will help determine how 
# many principal components we should use in our tSNE visualization

# Show this plot
screeplot(dat.filter.BCV.pca, npcs = 30, main = "All Data - High BCV Genes Screeplot")
abline(v=7,lwd=2,col="red")
ggscreeplot(dat.filter.BCV.pca)
# Conclustion: Seems to be clear that just the 
# first 6 PCs explain the most variation in our data
```

###Creating a t-SNE plot from the "significant" PCs with outliers removed
```{r tsne, warning=FALSE, message=FALSE}
nComponents<-6 # estimated from the screeplots
#seed<-runif(1,1,9999)
seed<-5084.043
set.seed(seed) #setting seed

dat.filter.BCV.tsne<-tsne(dat.filter.BCV.pca$x[,1:nComponents],perplexity=30,max_iter=2000,whiten = FALSE)

pData(dat.filter)$tSNE1_pos<-dat.filter.BCV.tsne[,1]
pData(dat.filter)$tSNE2_pos<-dat.filter.BCV.tsne[,2]

dat.filter.BCV.tsne.plot<-myTSNEPlotAlpha(dat.filter,color="region", shape="age") + scale_color_brewer(palette="Set1") + ggtitle("All valid cells - tSNE Plot")

dat.filter.BCV.tsne.plot
```

###Adding subset cluster annotation from recursive analyses
```{r adding subset clusters}
P7.Ob.clusters.df <- readRDS(file = file.path(Rdata_dir, "P7.Ob.clusters.df.rds"))
P7.Mb.clusters.df <- readRDS(file = file.path(Rdata_dir, "P7.Mb.clusters.df.rds"))
P7.Fb.clusters.df <- readRDS(file = file.path(Rdata_dir, "P7.Fb.clusters.df.rds"))
e15.Fb.clusters.df <- readRDS(file = file.path(Rdata_dir, "e15.Fb.clusters.df.rds"))
e15.Mb.clusters.df <- readRDS(file = file.path(Rdata_dir, "e15.Mb.clusters.df.rds"))

P7.Ob.clusters.df$subset.cluster <- paste0("P7.OB.",P7.Ob.clusters.df$kmeans_tSNE_cluster)
P7.Mb.clusters.df$subset.cluster <- paste0("P7.MB.",P7.Mb.clusters.df$kmeans_tSNE_cluster)
P7.Fb.clusters.df$subset.cluster <- paste0("P7.FB.",P7.Fb.clusters.df$kmeans_tSNE_cluster)

e15.Mb.clusters.df$subset.cluster <- paste0("E15.MB.",e15.Mb.clusters.df$kmeans_tSNE_cluster)
e15.Fb.clusters.df$subset.cluster <- paste0("E15.FB.",e15.Fb.clusters.df$kmeans_tSNE_cluster)

Sub.clusters.df <- rbind(P7.Ob.clusters.df,P7.Mb.clusters.df,P7.Fb.clusters.df,e15.Mb.clusters.df,e15.Fb.clusters.df)

names(Sub.clusters.df) <- c("sample_id.x","remove","subset.cluster")
final.data.frame <- merge(x = pData(dat.filter), y = Sub.clusters.df, by = "sample_id.x")
pData(dat.filter)$subset.cluster <- as.factor(final.data.frame$subset.cluster)

myTSNEPlotAlpha(dat.filter,color="subset.cluster", shape="age") + scale_color_brewer(palette="Set1") + ggtitle("All valid cells - tSNE Plot")
```

###Saving final all cell CDS for future analyses and figures
```{r saving analyzed cds for all}
dat.filter.final <- dat.filter
<<<<<<< HEAD
saveRDS(object = dat.filter.final, file = file.path(Rdata_dir, "dat.filter.final.Rds"), compress = "bzip2")
=======
saveRDS(object = dat.filter.final, file = file.path(Rdata_dir, "dat.filter.final.Rds"))
>>>>>>> 89c2d8fd5d2654e6a0382050ae9f801ae1bce1cf
```

###Session Info
```{r session infomation with outliers removed}
sessionInfo()
```

