---
title: "Cplx1 Striatum Th+ Density Analysis"
author: "Paul Hook"
output: html_document
editor_options: 
  chunk_output_type: console
---

***
**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`  

***

###Setting important directories. Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))

#loading any special libraries
library(nlme)
library(ggpubr)
library(ggbeeswarm)
library(DT)
```

###Loading data
First, we need to load the data from _Th_+ striatal quantification in WT and _Cplx1_ KO mice.
```{r loading}
# Loading data
dat <- read.delim(file = file.path(Rdata_dir,"combined_quantification.txt"))
# Making sure the 'mouse' column is a factor.
dat$mouse <- as.factor(dat$mouse)

datatable(data = dat, caption = "Th+ density measurements")
```


###Processing and inital visualization of the data
In order to analyze the data, we had to process it some first. "Blanks" which were measurements of the cortical _Th_+ staining in each section were removed, a unqiue identifier for each section was added and then mice were annotated as being "WT" or "KO"
```{r processing, fig.align='center'}
#Remove 0s 'adjusted.mean' which are from the cortical measurements
dat <- dat[dat$blank == "no",]

#adding a unique section variable
dat$section.sum <- paste0(dat$mouse, ".", dat$slide, ".", dat$section)

#KO and WT
ko <- c("xif.288", "xif.291", "xif.263","bif.299", "xim.932","xim.935","xif.122","xif.145")
wt <- c("xim.636","xim.666","xif.90")

#Annotating with KO or WT
dat <- dat %>%
  mutate(condition = factor(if_else(mouse %in% ko, "KO",
               if_else(mouse %in% wt, "WT", "NA")),levels = c("WT","KO")))

# Inital visualization of the data factored by condition
p <- ggplot(dat, aes(condition,adjusted.mean))
p + geom_violin() + geom_jitter(size = 2, width = 0.2) + ylim(c(0,125)) + ggtitle("Initial visualization of the data")
```

###Summarizing data
We decided to summarize the data by section since they were all independently stained and measured. This way we could also preserve variation seen within each mouse brain. The data summarized by mouse showed the same outcome. We also added information about age, date IHC was performed, and sex.
```{r summarizing data}
# Summarizing the data based on individual section means
dat.sum <- dat %>%
  dplyr::group_by(section.sum) %>%
  dplyr::summarize(section.mean = mean(adjusted.mean))

# Add in mouse annotation column
dat.sum$mouse <- gsub("^([^.]*.[^.]*)..*$", "\\1",dat.sum$section.sum)

# Add in condition and age annotation columns
dat.sum.new <- dat.sum %>%
  mutate(condition = factor(if_else(mouse %in% ko, "KO",
               if_else(mouse %in% wt, "WT", "NA")), levels = c("WT","KO"))) %>%
  mutate(age = factor(if_else(mouse %in% c("xif.90","bif.299","xim.932","xim.935","xif.122","xif.145"), "4",
                              if_else(mouse %in% c("xim.636","xif.288","xif.291", "xif.263"), "7",
                                      if_else(mouse %in% "xim.666", "7.5", "NA"))), levels = c("4","7",
                                                                                               "7.5"))) %>%
  mutate(date = factor(if_else(mouse %in% c("xim.636","xif.288","xif.291"), "4/20/04",
                              if_else(mouse %in% "xif.263", "12/02/04",
                                      if_else(mouse %in% "bif.299", "12/09/04",
                                              if_else(mouse %in% c("xif.90","xif.122","xif.145"), "5/19/05",
                                                      if_else(mouse %in% "xim.935", "6/12/05",
                                                              if_else(mouse %in% c("xim.932","xim.666"),
                                                                      "6/17/05", "NA")))))))) %>%
  mutate(sex = factor(if_else(mouse %in% c("xim.935", "xim.932","xim.666","xim.636"),"male","female")))
```

###Statistical Testing
We now wanted to test to see if there is a statistcally significant difference between _Th_+ fiber density in the striatum between _Cplx1_ KO and WT littermates. We do this by performing a two sided T-test.
```{r stats}
# First test if the variances in each condition are unequal
var <- var.test(section.mean~condition, dat.sum.new)
var

# Perform a two-sided t-test with equal variance
test <- t.test(section.mean~condition, dat.sum.new, alternative = "two.sided", var.equal = T, conf.level = 0.95)
test
```

###Visualization
```{r plot, fig.align='center',fig.height=4,fig.width=5}
# Plot the independent section measurements by condition
set.seed(2)
p <- ggplot(dat.sum.new, aes(condition,section.mean))
final <- p + geom_boxplot() + 
  geom_point(size = 1, position = position_jitterdodge(jitter.width = 0.5), aes(color = condition)) + 
  theme_pubr() + 
  ylab("Th+ striatum density\n(Th+ DAB Intensity/pixel squared)") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("blue","red")) + 
  theme(text = element_text(size=16)) + ylim(c(20,125)) +
  annotate("text", x=1.5,y=120,label=paste0("P = ",round(test$p.value,11)),size=4) +
  annotate("segment", x = 1, xend = 2, y = 115, yend = 115, lwd=0.5) +
  annotate("segment", x=1, xend=1,y=112,yend=115,lwd=0.5) +
  annotate("segment", x=2, xend=2,y=112,yend=115,lwd=0.5)

pdf(file = file.path(file_dir,"Figure.5E.pdf"), height = 4, width = 5)
final
dev.off()

final
```
