---
title: "E15 MB Analysis"
author: "Paul Hook"
output: html_document
---

**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

**Setting important directories. Also loading important libraries and custom functions for analysis.**
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))
```

**Loading the E15 cds**
```{r Loading .rds}
e15.Mb.dat.filter <- readRDS(file = file.path(Rdata_dir, "e15.Mb.dat.filter.rds"))
```

**Filter genes by percentage of cells expresssing each gene**
```{r filter by cells expressing each gene}
# Plot number of cells expressing each gene as histogram
hist(fData(e15.Mb.dat.filter)$num_cells_expressed,breaks=50,col="red",main="Cells expressed per gene")

# Keep only expressed genes with expression in >= 5% of cells
numCellThreshold<-nrow(pData(e15.Mb.dat.filter))*0.05
e15.Mb.dat.expressed_genes<-row.names(subset(fData(e15.Mb.dat.filter),num_cells_expressed >= numCellThreshold))

# Same plot as above with threshold
hist(fData(e15.Mb.dat.filter)$num_cells_expressed,breaks=50,col="red",main="Cells expressed per gene - threshold")
abline(v=numCellThreshold,lty="dashed")
```

**Prepping the Monocole model for analysis**
```{r model_prep - Preparing the data for monocle analysis}
# Only keeping "expressed" genes
e15.Mb.dat.filter <-e15.Mb.dat.filter[e15.Mb.dat.expressed_genes,]

# Estimating the size factors
e15.Mb.dat.filter <-estimateSizeFactors(e15.Mb.dat.filter)

# Estimating dispersions
e15.Mb.dat.filter <- estimateDispersions(e15.Mb.dat.filter,cores=8)
# Removing 134 outliers
# Warning message:
# Deprecated, use tibble::rownames_to_column() instead. 
```

**Calculating summary stats**
```{r summary_stats}
# Calculating summary stats
fData(e15.Mb.dat.filter)$mean_expr<-apply(round(exprs(e15.Mb.dat.filter)),1,mean) # mean expression
fData(e15.Mb.dat.filter)$sd_expr<-apply(round(exprs(e15.Mb.dat.filter)),1,sd) # sd expression
fData(e15.Mb.dat.filter)$bcv<-(fData(e15.Mb.dat.filter)$sd_expr/fData(e15.Mb.dat.filter)$mean_expr)**2 # calculating biological coefficient of variation
fData(e15.Mb.dat.filter)$percent_detection<-(fData(e15.Mb.dat.filter)$num_cells_expressed/dim(e15.Mb.dat.filter)[2])*100 # calculating % detection
```

**Identifying high dispersion genes**
```{r high_dispersion_genes_monocle - Pulling out the high dispersion genes for PCA analysis}
e15.Mb.dat.filter.genes <- e15.Mb.dat.filter # spoofing the CellDataSet
disp_table <- dispersionTable(e15.Mb.dat.filter.genes) # pulling out the dispersion table
unsup_clustering_genes <-subset(disp_table, mean_expression >= 0.5 & dispersion_empirical >= 2 * dispersion_fit) # subsetting the data to pull out genes with expression above 0.5 and dispersion empirical > 2
e15.Mb.dat.high_bcv_genes<-unsup_clustering_genes$gene_id # pulling out list of genes
e15.Mb.dat.filter.order <- setOrderingFilter(e15.Mb.dat.filter, unsup_clustering_genes$gene_id)
plot_ordering_genes(e15.Mb.dat.filter.order) # plotting the dispersion and genes
length(e15.Mb.dat.high_bcv_genes) # 778
```

**Running PCA with high dispersion genes**
```{r Run PCA with high BCV genes}
# BCV Identified high dispersion genes. Running PC analysis
e15.Mb.dat.filter.BCV.pca<-prcomp(t(log2(exprs(e15.Mb.dat.filter[e15.Mb.dat.high_bcv_genes,])+1)),center=T,scale. = TRUE)

# Plotting the PCA graphs
# Plotting the first 2 PCs and coloring by age
hvPCA1<-ggbiplot(e15.Mb.dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(e15.Mb.dat.filter)$age,ellipse=T,var.axes=F) + scale_color_manual(values=c("darkgreen","red")) + monocle:::monocle_theme_opts()

# Plotting the first 2 PCs and coloring by region
hvPCA2<-ggbiplot(e15.Mb.dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(e15.Mb.dat.filter)$region,ellipse=T,var.axes=F) + scale_color_brewer(palette="Set1") + monocle:::monocle_theme_opts()

# Plotting the first 2 PCs and coloring by plate the cell was sequenced from
hvPCA3<-ggbiplot(e15.Mb.dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(e15.Mb.dat.filter)$split_plate,ellipse=T,var.axes=F) + scale_color_brewer(palette="Set1") + monocle:::monocle_theme_opts()

# Show the plots in the terminal
hvPCA1
hvPCA2
hvPCA3
```

**Viewing screeplots and determining the number of "significant" PCs**
```{r Screeplots of Data,tidy=TRUE}
# Making a screeplot of the BCV PCA. This will help determine how 
# many principal components we should use in our tSNE visualization

# Show this plot

screeplot(e15.Mb.dat.filter.BCV.pca, npcs = 30, main = "e15.5 - High BCV Genes Screeplot")
abline(v=5, lwd=2, col="red")
ggscreeplot(e15.Mb.dat.filter.BCV.pca)

# Conclustion: Seems to be clear that just the 
# first four PCs explain the most variation in our data
```

**Creating a t-SNE plot from the "significant" PCs**
```{r tsne, message=FALSE, warning=FALSE}
nComponents<-4 # estimated from the screeplots
#seed <- runif(1,1,9999) # determined by testing random seeds
seed <- 223.3888
set.seed(seed) #setting seed

e15.Mb.dat.filter.BCV.tsne<-tsne(e15.Mb.dat.filter.BCV.pca$x[,1:nComponents],perplexity=20,max_iter=5000,whiten = FALSE)

pData(e15.Mb.dat.filter)$tSNE1_pos<-e15.Mb.dat.filter.BCV.tsne[,1]
pData(e15.Mb.dat.filter)$tSNE2_pos<-e15.Mb.dat.filter.BCV.tsne[,2]

e15.Mb.dat.filter.BCV.tsne.plot<-myTSNEPlotAlpha(e15.Mb.dat.filter,color="region", shape="age") + scale_color_brewer(palette="Set1") + ggtitle("e15.5 - BCV tSNE Plot")

e15.Mb.dat.filter.BCV.tsne.plot
```

**Identifying clusters in the data in an unsupervised manner**
```{r - attempting to cluster,tidy=TRUE}
# Going to attempt to use the R program "ADPclust" to determine how many clusters our data has

# Loading NbClust
library(ADPclust)

# Running ADPclust
clust.res <- adpclust(x = e15.Mb.dat.filter.BCV.tsne, draw = T)

# Extracting the "best partition" (aka the best cluster) for each cell
clust.res.df <- as.data.frame(clust.res$cluster)

# Adding the cluster assignment for each cell to the pData
pData(e15.Mb.dat.filter)$kmeans_tSNE_cluster <- as.factor(clust.res.df$`clust.res$cluster`)

# Plotting the same tSNE plot as above but coloring with the "clusters"
myTSNEPlotAlpha(e15.Mb.dat.filter,color="kmeans_tSNE_cluster", shape="age") + scale_color_brewer(palette="Set1") + ggtitle("All Cells - BCV tSNE with Clusters Plot")
```

**QC on the clusters**
```{r check the clusters - qc on clusters}
# Distribution of number of genes expressed in clusters
q <- ggplot(pData(e15.Mb.dat.filter)) +
  geom_density(aes(x=num_genes_expressed,fill=kmeans_tSNE_cluster),alpha=0.3) + scale_fill_brewer(palette="Set1") + facet_grid(.~age) + monocle:::monocle_theme_opts()

q

# Plotting the distribution of total mRNAs in clusters
q<-ggplot(pData(e15.Mb.dat.filter)) +
  geom_density(aes(x=Total_mRNAs,fill=kmeans_tSNE_cluster),alpha=0.3) + scale_fill_brewer(palette="Set1") + facet_grid(.~age) + monocle:::monocle_theme_opts()

q
```

**Saving the E15.5 MB final cds**
```{r saving this .Rdata so it can be loaded in various scripts }
e15.Mb.dat.filter.final <- e15.Mb.dat.filter
saveRDS(object = e15.Mb.dat.filter.final, file = file.path(Rdata_dir, "e15.Mb.dat.filter.final.Rds"))
```

**Extracting pData information and saving it**
```{r extacting pData information and subset cluster information}
e15.Mb.clusters.df <- pData(e15.Mb.dat.filter)[,c(1,39)]
saveRDS(e15.Mb.clusters.df, file = file.path(Rdata_dir, "e15.Mb.clusters.df.rds"))
```

###Session Info
```{r session infomation}
sessionInfo()
```