---
title: "Supplemental Figure 1"
author: "Paul Hook"
output: html_document
---
***
**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`  

***

###Setting important directories. Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))

#loading any special libraries
library(readr)
library(cowplot)
```

### Loading the data needed
The final 'dat' cds was loaded and some summary stats were calculated to quickly check the data
```{r loading data needed}
# Filtered final cds
dat.filter <- readRDS(file.path(Rdata_dir,"dat.filter.final.Rds"))
nrow(pData(dat.filter)) # 396
with(pData(dat.filter), table(age, region))

# Outlier plots
FB.outlier.plot <- readRDS(file = file.path(Rdata_dir,"P7.FB.PCA.outliers.rds"))
MB.outlier.plot <- readRDS(file = file.path(Rdata_dir,"P7.MB.PCA.outliers.rds"))
OB.outlier.plot <- readRDS(file = file.path(Rdata_dir,"P7.OB.PCA.outliers.rds"))
```

###Figure S1A
Histogram displaying number of genes per cell in the final filtered dataset.
```{r Generating QC plots, fig.align='center',fig.height=6,fig.width=6}
# Number of genes per cell
max.expressed <- round(max(pData(dat.filter)$num_genes_expressed))
min.expressed <- round(min(pData(dat.filter)$num_genes_expressed))
mean.expressed <- round(mean(pData(dat.filter)$num_genes_expressed))

# Writing out histogram to pdf
pdf(file = file.path(file_dir, "Figure.S1A.pdf"), width = 6, height = 6)
hist(pData(dat.filter)$num_genes_expressed,
     breaks=25,
     main = NA,
     col="steelblue",
     xlab = "Number of genes expressed",
     ylab = "Number of cells",
     ylim = c(0,30))
mtext(text = paste("Min =",min.expressed), adj = 0.15, padj = 10)
mtext(text = paste("Max =", max.expressed), adj = 0.15, padj = 12)
mtext(text = paste("Mean =",mean.expressed), adj = 0.15, padj = 14)
title("Genes expressed per cell", line = -0.3)
dev.off()

# Plotting histogram
hist(pData(dat.filter)$num_genes_expressed,
     breaks=25,
     main = NA,
     col="steelblue",
     xlab = "Number of genes expressed",
     ylab = "Number of cells",
     ylim = c(0,30))
mtext(text = paste("Min =",min.expressed), adj = 0.15, padj = 10)
mtext(text = paste("Max =", max.expressed), adj = 0.15, padj = 12)
mtext(text = paste("Mean =",mean.expressed), adj = 0.15, padj = 14)
title("Genes expressed per cell", line = -0.3)
```

###Figure S1B
Histogram displaying mRNA per cell in the final filtered dataset.
```{r, fig.align='center',fig.height=6, fig.width=6}
# mRNA per cell
max.mRNAs <- round(max(pData(dat.filter)$Total_mRNAs))
min.mRNAs <- round(min(pData(dat.filter)$Total_mRNAs))
mean.mRNAs <- round(mean(pData(dat.filter)$Total_mRNAs))

# Writing out histograms to pdf
pdf(file = file.path(file_dir, "Figure.S1B.pdf"), width = 6, height = 6)
hist(pData(dat.filter)$Total_mRNAs,breaks=50,
     col="darkgreen",
     main = NA,
     ylab = "Number of cells", xlab = "Total mRNAs",
     ylim = c(0,25))
mtext(text = paste("Min =",min.mRNAs), adj = 0.15, padj = 10)
mtext(text = paste("Max =", max.mRNAs), adj = 0.15, padj = 12)
mtext(text = paste("Mean =",mean.mRNAs), adj = 0.15, padj = 14)
title("Total mRNA per cell", line = -0.3)
dev.off()

# Plotting the histogram
hist(pData(dat.filter)$Total_mRNAs,breaks=50,
     col="darkgreen",
     main = NA,
     ylab = "Number of cells", xlab = "Total mRNAs",
     ylim = c(0,25))
mtext(text = paste("Min =",min.mRNAs), adj = 0.15, padj = 10)
mtext(text = paste("Max =", max.mRNAs), adj = 0.15, padj = 12)
mtext(text = paste("Mean =",mean.mRNAs), adj = 0.15, padj = 14)
title("Total mRNA per cell", line = -0.3)
```

###Figure S1C
Histogram displaying mass per cell in the final filtered dataset.
```{r,fig.align='center',fig.height=6, fig.width=6}
# Mass per cell
max.mass <- round(max(pData(dat.filter)$total_mass))
min.mass <- round(min(pData(dat.filter)$total_mass))
mean.mass <- round(mean(pData(dat.filter)$total_mass))

# Writing out the histogram to a PDF
pdf(file = file.path(file_dir, "Figure.S1C.pdf"), width = 6, height = 6)
hist(pData(dat.filter)$total_mass,
     breaks=50,
     main = NA,
     col="darkred",
     ylab = "Number of cells", xlab = "Total mass\n(fragments mapped to transcriptome)",
     ylim = c(0,20))
mtext(text = paste("Min =",min.mass), adj = 0.85, padj = 10)
mtext(text = paste("Max =", max.mass), adj = 0.85, padj = 12)
mtext(text = paste("Mean =",mean.mass), adj = 0.85, padj = 14)
title("Total mass per cell", line = -0.3)
dev.off()

# Plotting the histogram
hist(pData(dat.filter)$total_mass,
     breaks=50,
     main = NA,
     col="darkred",
     ylab = "Number of cells", xlab = "Total mass\n(fragments mapped to transcriptome)",
     ylim = c(0,20))
mtext(text = paste("Min =",min.mass), adj = 0.85, padj = 10)
mtext(text = paste("Max =", max.mass), adj = 0.85, padj = 12)
mtext(text = paste("Mean =",mean.mass), adj = 0.85, padj = 14)
title("Total mass per cell", line = -0.3)
```

###Figure S1D
Barplots showing the number of final cells by condition assayed after filtering
```{r, fig.align='center',fig.height=4,fig.width=6}
# Number of cells x condition
p <- ggplot(pData(dat.filter))
q <- p + geom_bar(aes(x=region,fill=region),width=0.5) + ggtitle("Valid cells per condition") + theme_bw() + scale_fill_brewer(palette="Set1", guide = F) + facet_grid(.~age) + ylab("Number of cells") + xlab("Region") + theme(plot.title = element_text(hjust = 0.5))

pdf(file = file.path(file_dir, "Figure.S1D.pdf"), width = 6, height = 4)
q
dev.off()

q
```

###Figure S1E
PCA plots displaying outliers that were removed during data filtering. Note that circles appearing on the plots were removed.
```{r, fig.align='center',fig.height=3,fig.width=13}
fb <- FB.outlier.plot + scale_color_manual(values = brewer.pal(8,"Set1")[1], name = "Region") + monocle:::monocle_theme_opts() + ggtitle("P7 FB PCA (PC1 vs. PC2)") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(family = 'Helvetica',size=10),
        axis.title.x = element_text(family = 'Helvetica',size=10))

mb <- MB.outlier.plot + scale_color_manual(values = brewer.pal(8,"Set1")[2], name = "Region") + monocle:::monocle_theme_opts() + ggtitle("P7 MB PCA (PC1 vs. PC2)") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(family = 'Helvetica',size=10),
        axis.title.x = element_text(family = 'Helvetica',size=10))

ob <- OB.outlier.plot + scale_color_manual(values = brewer.pal(8,"Set1")[3], name = "Region") + monocle:::monocle_theme_opts() + ggtitle("P7 OB PCA (PC1 vs. PC2)") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(family = 'Helvetica',size=10),
        axis.title.x = element_text(family = 'Helvetica',size=10))

pdf(file = file.path(file_dir, "Figure.S1E.pdf"), width = 4, height = 3)
fb
ob
mb
dev.off()

plot_grid(fb,ob,mb,nrow = 1)
```

###Session Info
```{r session info}
sessionInfo()
```