---
title: "Figure.S4.Rmd"
author: "Paul Hook"
output: html_document
editor_options: 
  chunk_output_type: console
---

***
**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`  

***

###Setting important directories. 
Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))

#loading any special libraries
library(readr)
library(cowplot)
```

###Setting a plot theme
```{r, setting a theme}
theme <- theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(family="Helvetica",size=8),
        axis.text.y = element_text(size=6),
        strip.text = element_text(family="Helvetica",face="bold.italic",size=10))

```

###Loading the data needed
```{r loading data needed,message=FALSE,warning=FALSE}
###Loading the cds data needed to produce figures
dat.filter <- readRDS(file.path(Rdata_dir,"dat.filter.final.Rds"))
color <- c("#F6937A","#882F1C","#BF1F36","#C2B280","#F2C318","#E790AC","#F18421","#0168A5","#848483","#A4CAEB","#885793","#008957","#222222")
```

###Figure S4A
This panel consisted for various boxplots related to canonical markers of neuronal subtypes including dopaminergic, GABAergic, and glutamatergic neurons. Labels seen in Figure S4A were added later.
```{r FB Making boxplots, fig.align='center',fig.width=8,fig.height=8}
th <- myBoxplot.subset(dat.filter, "Th", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme 

gad1 <- myBoxplot.subset(dat.filter, "Gad1", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme + theme(axis.title.y=element_blank())

slc6a3 <- myBoxplot.subset(dat.filter, "Slc6a3", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme 

gad2 <- myBoxplot.subset(dat.filter, "Gad2", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme + theme(axis.title.y=element_blank())

slc18a2 <- myBoxplot.subset(dat.filter, "Slc18a2", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme

slc32a1 <- myBoxplot.subset(dat.filter, "Slc32a1", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme + theme(axis.title.y=element_blank()) + scale_y_continuous(breaks = c(0,1,2,3,4))

ddc <- myBoxplot.subset(dat.filter, "Ddc", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme

#Make blank plot
df <- data.frame()
blank <- ggplot(df) + geom_point() + xlim(0, 10) + ylim(0, 100) + geom_blank() + theme_nothing()

egfp <- myBoxplot.subset(dat.filter, "eGFP", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme

slc17a6 <- myBoxplot.subset(dat.filter, "Slc17a6", logMode = T) + scale_fill_manual(values = color, name = "Subset Cluster") + theme + theme(axis.title.y=element_blank())

grid_arrange_shared_legend(th,gad1,slc6a3,gad2,slc18a2,slc32a1,ddc,blank,egfp,slc17a6,ncol=2,nrow=5, position = "right")

pdf(file = file.path(file_dir, "Figure.S4A.pdf"), height = 2, width = 4)
th
slc6a3
slc18a2
ddc
egfp
gad1
gad2
slc32a1
slc17a6
dev.off()
```

###Session Info
```{r session info}
sessionInfo()
```
