---
title: "Figure 2"
author: "Paul Hook"
output: html_document
editor_options: 
  chunk_output_type: console
---
***
**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`  

***

###Setting important directories. Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
figure_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/figures/"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))
```

###Loading the cds data needed to produce figures
```{r loading data needed}
dat.filter <- readRDS(file.path(Rdata_dir,"dat.filter.final.Rds"))
P7.dat.filter <- readRDS(file.path(Rdata_dir,"P7.dat.filter.final.Rds"))
P7.MB.dat.filter <- readRDS(file.path(Rdata_dir,"P7.Mb.dat.filter.final.Rds"))
P7.FB.dat.filter <- readRDS(file.path(Rdata_dir,"P7.Fb.dat.filter.final.Rds"))
P7.OB.dat.filter <- readRDS(file.path(Rdata_dir,"P7.Ob.dat.filter.final.Rds"))
```

### Figure 2A
The following code was used to produce Figure 2A, the P7 all cell t-SNE plot colored by subset cluster identity
```{r making P7 t-SNE plots, Figure 2A,fig.align='center',fig.width=6,fig.height=6}
# Seting the subset cluster colors needed to keep everything consistent
color.P7 <- c("#F2C318","#E790AC","#F18421","#0168A5","#848483","#A4CAEB","#885793","#008957","#222222")

# Plotting the t-SNE of P7 cells colored by subset cluster identity
p7 <- myTSNEPlotAlpha.triangle(P7.dat.filter,color="subset.cluster") + scale_color_manual(values = color.P7, name = "Subset Cluster") + ggtitle("P7 neurons - tSNE Plot by subset cluster") + xlab("tSNE 1") + ylab("tSNE 2") + theme(legend.position = "bottom",axis.line = element_line(colour = "black")) + theme(plot.title = element_text(hjust = 0.5)) + coord_fixed(ratio = 1) + scale_shape(name = "Age")

pdf(file = file.path(figure_dir, "Figure.2A.pdf"), width = 8, height = 8)
p7
dev.off()

p7
```

###Figure 2B
The following code was used to produce Figure 2B, 2C, and 2D. These figures display the t-SNE results for the regional P7 analyses including FB, OB, and MB.
```{r ob tsne, fig.align='center',fig.width=4,fig.height=4}
# Plotting the t-SNE of P7 OB cells colored by the clusters identified during the iterative analysis
ob <- myTSNEPlotAlpha.triangle(P7.OB.dat.filter,color="kmeans_tSNE_cluster") + scale_color_manual(values = color.P7[7:9], name = "Subset Cluster", labels = c("P7.OB.1","P7.OB.2","P7.OB.3")) + ggtitle("P7 OB neurons t-SNE") + xlab("t-SNE 1") + ylab("t-SNE 2") + theme(legend.position = "bottom",axis.line = element_line(colour = "black")) + theme(plot.title = element_text(hjust = 0.5)) + coord_fixed(ratio = 1) + scale_shape(name = "Age")

pdf(file.path(figure_dir,"Figure.2C.pdf"), height = 4, width = 4)
ob
dev.off()

ob
```

```{r fb tsne ,fig.align='center',fig.height=5,fig.width=5}
# Plotting the t-SNE of P7 FB cells colored by the clusters identified during the iterative analysis
fb <- myTSNEPlotAlpha.triangle(P7.FB.dat.filter,color="kmeans_tSNE_cluster") + scale_color_manual(values = color.P7[1:2], name = "Subset Cluster", labels = c("P7.FB.1","P7.FB.2")) + ggtitle("P7 FB t-SNE") + xlab("t-SNE 1") + ylab("t-SNE 2") + theme(legend.position = "bottom",axis.line = element_line(colour = "black")) + theme(plot.title = element_text(hjust = 0.5)) + coord_fixed(ratio = 1) + scale_shape(name = "Age") + coord_equal(ratio = 0.5)

pdf(file.path(figure_dir,"Figure.2B.pdf"), height = 5, width = 5)
fb
dev.off()

fb
```

```{r mb tsne ,fig.align='center',fig.height=6,fig.width=5}
# Plotting the t-SNE of P7 MB cells colored by the clusters identified during the iterative analysis
mb <- myTSNEPlotAlpha.triangle(P7.MB.dat.filter,color="kmeans_tSNE_cluster") + scale_color_manual(values = color.P7[3:6], name = "Subset Cluster", labels = c("P7.MB.1","P7.MB.2","P7.MB.3","P7.MB.4")) + ggtitle("P7 MB t-SNE") + xlab("t-SNE 1") + ylab("t-SNE 2") + theme(legend.position = "bottom",axis.line = element_line(colour = "black")) + theme(plot.title = element_text(hjust = 0.5)) + coord_fixed(ratio = 0.5) + scale_shape(name = "Age")

pdf(file = file.path(figure_dir, "Figure.2D.pdf"), width = 6, height = 4)
mb
dev.off()

mb
```

###Session Info
```{r session info}
sessionInfo()
```
