---
title: "P7 All Recursive Analysis"
author: "Paul Hook"
output: html_document
---

**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

###Setting important directories
Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))
```

###Loading P7 all cells .rds
```{r loading .rds}
P7.dat.filter <- readRDS(file = file.path(Rdata_dir, "P7.dat.filter.rds"))
```

###Removing outliers identified in the recursive analyses
```{r Remove outliers}
# Removing outliers from other analyses
Ob.outliers <- readRDS(file = file.path(Rdata_dir, "P7.Ob.outliers.rds"))

Mb.outliers <- readRDS(file = file.path(Rdata_dir, "P7.Mb.outliers.rds"))

Fb.outliers <- readRDS(file = file.path(Rdata_dir, "P7.Fb.outliers.rds"))

P7.outliers <- c(Mb.outliers, Ob.outliers, Fb.outliers)

saveRDS(object = P7.outliers, file = file.path(Rdata_dir,"P7.outliers.rds"))

P7.dat.filter <- readRDS(file = file.path(Rdata_dir, "P7.dat.filter.rds"))
nrow(pData(P7.dat.filter)) # 238

P7.dat.filter <- P7.dat.filter[,!(pData(P7.dat.filter)$sample_id.x %in% P7.outliers)]

nrow(pData(P7.dat.filter)) #224
```

###Filter genes by percentage of cells expresssing each gene
```{r Filter by cells per genes}
# Plot number of cells expressing each gene as histogram
hist(fData(P7.dat.filter)$num_cells_expressed,breaks=50,col="red",main="Cells expressed per gene")

# Keep only expressed genes with expression in >= 5% of cells
numCellThreshold<-nrow(pData(P7.dat.filter))*0.05
P7.dat.expressed_genes<-row.names(subset(fData(P7.dat.filter),num_cells_expressed >= numCellThreshold))

# Same plot as above with threshold
hist(fData(P7.dat.filter)$num_cells_expressed,breaks=50,col="red",main="Cells expressed per gene - threshold")
abline(v=numCellThreshold,lty="dashed")
```

###Prepping the Monocle model for analysis
```{r model_prep - Preparing the data for monocle analysis}
# Only keeping "expressed" genes
P7.dat.filter <-P7.dat.filter[P7.dat.expressed_genes,]

# Estimating the size factors
P7.dat.filter <-estimateSizeFactors(P7.dat.filter)

# Estimating dispersions
P7.dat.filter <- estimateDispersions(P7.dat.filter,cores=8)
# Removing 229 outliers
# Warning message:
# Deprecated, use tibble::rownames_to_column() instead. 
```

###Calculating summary stats
```{r summary_stats}
# Calculating summary stats
fData(P7.dat.filter)$mean_expr<-apply(round(exprs(P7.dat.filter)),1,mean) # mean expression
fData(P7.dat.filter)$sd_expr<-apply(round(exprs(P7.dat.filter)),1,sd) # sd expression
fData(P7.dat.filter)$bcv<-(fData(P7.dat.filter)$sd_expr/fData(P7.dat.filter)$mean_expr)**2 # calculating biological coefficient of variation
fData(P7.dat.filter)$percent_detection<-(fData(P7.dat.filter)$num_cells_expressed/dim(P7.dat.filter)[2])*100 # calculating % detection
```


###Identifying high dispersion genes
```{r high_dispersion_genes_monocle - Pulling out the high dispersion genes for PCA analysis}
P7.dat.filter.genes <- P7.dat.filter # spoofing the CellDataSet
disp_table <- dispersionTable(P7.dat.filter.genes) # pulling out the dispersion table
unsup_clustering_genes <-subset(disp_table, mean_expression >= 0.5 & dispersion_empirical >= 1.5 * dispersion_fit) # subsetting the data to pull out genes with expression above 0.5 and dispersion empirical > 2
P7.dat.high_bcv_genes<-unsup_clustering_genes$gene_id # pulling out list of genes
P7.dat.filter.order <- setOrderingFilter(P7.dat.filter, unsup_clustering_genes$gene_id)
plot_ordering_genes(P7.dat.filter.order) # plotting the dispersion and genes
length(P7.dat.high_bcv_genes) # 1120
```


###Running PCA with high dispersion genes
```{r Run PCA with high BCV genes}

# BCV Identified high dispersion genes. Running PC analysis
P7.dat.filter.BCV.pca<-prcomp(t(log2(exprs(P7.dat.filter[P7.dat.high_bcv_genes,])+1)),center=T,scale. = TRUE)

# Plotting the PCA graphs
# Plotting the first 2 PCs and coloring by age
hvPCA1<-ggbiplot(P7.dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(P7.dat.filter)$age,ellipse=T,var.axes=F) + scale_color_manual(values=c("darkgreen","red")) + monocle:::monocle_theme_opts()

# Plotting the first 2 PCs and coloring by region
hvPCA2<-ggbiplot(P7.dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(P7.dat.filter)$region,ellipse=T,var.axes=F) + scale_color_brewer(palette="Set1") + monocle:::monocle_theme_opts() + ggtitle("All P7 PCA")

# Plotting the first 2 PCs and coloring by plate the cell was sequenced from
hvPCA3<-ggbiplot(P7.dat.filter.BCV.pca,choices=c(1,2),scale=0,groups=pData(P7.dat.filter)$split_plate,ellipse=T,var.axes=F) + scale_color_brewer(palette="Set1") + monocle:::monocle_theme_opts()

# Show the plots in the terminal
hvPCA1
hvPCA2
hvPCA3

# No real outliers seen
```


###Screeplots
Viewing screeplots and determining the number of "significant" PCs.
```{r Screeplots of Data,tidy=TRUE}
# Making a screeplot of the BCV PCA. This will help determine how 
# many principal components we should use in our tSNE visualization

# Show this plot
screeplot(P7.dat.filter.BCV.pca, npcs = 30, main = "e15.5 - High BCV Genes Screeplot")
abline(v=5,lwd=2,col="red")
ggscreeplot(P7.dat.filter.BCV.pca)

# Conclustion: Seems to be clear that just the 
# first 4 PCs explain the most variation in our data
```

###Creating a t-SNE plot from the "significant" PCs with outliers removed
```{r tsne, warning=FALSE,message=FALSE}
nComponents<-4 # estimated from the screeplots
#seed <- runif(1,1,9999) # determined by testing random seeds
seed <- 1974.079
set.seed(seed) #setting seed

P7.dat.filter.BCV.tsne<-tsne(P7.dat.filter.BCV.pca$x[,1:nComponents],perplexity=30,max_iter=3000,whiten = FALSE)

pData(P7.dat.filter)$tSNE1_pos<-P7.dat.filter.BCV.tsne[,1]
pData(P7.dat.filter)$tSNE2_pos<-P7.dat.filter.BCV.tsne[,2]

P7.dat.filter.BCV.tsne.plot<-myTSNEPlotAlpha(P7.dat.filter,color="region", shape="age") + scale_color_brewer(palette="Set1") + ggtitle("P7 - BCV tSNE Plot")

P7.dat.filter.BCV.tsne.plot
```

###Adding MB/FB/OB subcluster annotation
```{r Adding the MB/FB/OB subclusters}
P7.Ob.clusters.df <- readRDS(file = file.path(Rdata_dir, "P7.Ob.clusters.df.rds"))
P7.Mb.clusters.df <- readRDS(file = file.path(Rdata_dir, "P7.Mb.clusters.df.rds"))
P7.Fb.clusters.df <- readRDS(file = file.path(Rdata_dir, "P7.Fb.clusters.df.rds"))

P7.Ob.clusters.df$subset.cluster <- paste0("P7.OB.",P7.Ob.clusters.df$kmeans_tSNE_cluster)
P7.Mb.clusters.df$subset.cluster <- paste0("P7.MB.",P7.Mb.clusters.df$kmeans_tSNE_cluster)
P7.Fb.clusters.df$subset.cluster <- paste0("P7.FB.",P7.Fb.clusters.df$kmeans_tSNE_cluster)

Sub.clusters.df <- rbind(P7.Ob.clusters.df,P7.Mb.clusters.df,P7.Fb.clusters.df)

names(Sub.clusters.df) <- c("sample_id.x","remove","subset.cluster")
final.data.frame <- merge(x = pData(P7.dat.filter), y = Sub.clusters.df, by = "sample_id.x")
pData(P7.dat.filter)$subset.cluster <- as.factor(final.data.frame$subset.cluster)

myTSNEPlotAlpha(P7.dat.filter,color="subset.cluster", shape="age") + scale_color_brewer(palette="Set1") + ggtitle("P7 - BCV tSNE Plot")
```

###Saving the P7 final cds 
```{r saving analyzed cds for P7 all}
P7.dat.filter.final <- P7.dat.filter
saveRDS(object = P7.dat.filter.final, file = file.path(Rdata_dir, "P7.dat.filter.final.Rds"))
```

###Session Info
```{r session infomation with outliers removed}
sessionInfo()
```

