---
title: "P7 WGCNA Analysis"
author: "Paul Hook"
output: html_document
---

**Last update:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

###Setting important directories
Also loading important libraries and custom functions for analysis.
```{r init, message=FALSE, warning=FALSE}
seq_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
file_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/output"
Rdata_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/data"
Script_dir <- "/Volumes/PAULHOOK/sc-da-parkinsons/code"
source(file.path(Script_dir,'init.R'))
source(file.path(Script_dir,"tools_R.r"))
```

### Loading R libraries necessary for WGCNA analysis
```{r Load necessary libraries, message=FALSE, warning=FALSE}
library(AnnotationDbi)
library(impute)
library(GO.db)
library(preprocessCore)
library(knitr)
library(rgl)
library(WGCNA)
library(SC3)
library(clusterProfiler)
library(org.Mm.eg.db)
library(ReactomePA)
library(cummeRbund)
library(foreach)
library(doParallel)
library(splines)
library(flashClust)
```

### Loading in needed P7 data
```{r Loading data}
P7.dat.filter <- readRDS(file.path(Rdata_dir,"P7.dat.filter.final.Rds"))
dim(P7.dat.filter)
```

### Preparing data for WGCNA analysis
Preparing the data for WGCNA analysis. The genes used in the analysis were filtered to only include genes that were determined to be expressed in at least 20 cells (9% of the total cells in the data set).
```{r}
# Identifying genes only expressed in at least 20 cells
P7.expressed_genes.WGCNA<-row.names(subset(fData(P7.dat.filter),num_cells_expressed >= 20))
length(P7.expressed_genes.WGCNA)

# Keeping only those genes for analysis. Ended up being 12,628 cells.
P7.dat.WGCNA <-P7.dat.filter[P7.expressed_genes.WGCNA,]

# Converting data to log2 + 1 expression
mydata <- log2(exprs(P7.dat.WGCNA)+1)

# Converting data to a matrix which is needed for WGCNA analysis.
mydata <- as.matrix(mydata)

# Checking the dimensions of the data
dim(mydata)

# Transposing the matrix so it is in the orientation needed for WGCNA analysis and saving the matrix under a better variable name.
mydata.trans=t(mydata)
datExpr <- mydata.trans
```

### Heirarchial clustering of cells
Next, heirarchial clutering was used to cluster the data and look for any outliers. There seems to be only one glaring outlier.
```{r Clustering cells and plotting, cache=TRUE}
# Clustering the data by using heirarchial clustering
sampleTree <- flashClust(dist(datExpr), method = "average")

# Plotting the heirarchial clustering in order to manually identify and removce outliers
par(cex = 0.6, mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers",
     sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
cutHeight=120 # choosing a cut height of 120 in order to remove the one high level outlier
abline(h = cutHeight, col = "red")

# Cutting the tree at the cut height and seeing how many outliers are removed with the cut
clust = cutreeStatic(sampleTree, cutHeight = cutHeight, minSize = 10)
table(clust)

# Keeping only the samples that are not removed by the cut height used
keepSamples <- (clust==1)
datExpr <- datExpr[keepSamples,]

# Saving the number of genes and the number of samples that will move in to WGCNA
nGenes <- ncol(datExpr)
nSamples <- nrow(datExpr)
```

### Making simple trait dataframe
Making a data frame that can be filled with meta-data about the cells and can be used to annotate the heirarchial clustering
```{r Making trait dataframe, message=FALSE, warning=FALSE}
# subsetting out trait data from the initial cds
datTraits <- pData(P7.dat.WGCNA[,keepSamples])
datTraits<-datTraits[,c("region","subset.cluster","num_genes_expressed")]

# Setting the colors to match the regional colors
region_labels <- RColorBrewer::brewer.pal(6, "Set1")[1:3]

# Making sure the regional colors match what has been used througout and making a column with the hex codes.
datTraits <- datTraits %>%
  mutate(cell.names = row.names(datTraits)) %>%
  mutate(x = ifelse(datTraits$region == "FB", region_labels[1],
                    ifelse(datTraits$region == "MB", region_labels[2],
                           ifelse(datTraits$region == "OB", region_labels[3], NA))))
  
# Re-performing heirarchial clustering
sampleTree2 <- hclust(dist(datExpr), method = "average")

# Writing out that dendrogram to a PDF
pdf(file = file.path(file_dir, "sample.dendrogram_region.heatmap.pdf"), height = 8, width = 12)
plotDendroAndColors(sampleTree2, datTraits$x,
                groupLabels = names(datTraits),
                main = "Sample dendrogram and regional identity")
dev.off()

# Plotting the dendrogram with regional identity colored beneath each cell
plotDendroAndColors(sampleTree2, datTraits$x,
                groupLabels = names(datTraits),
                main = "Sample dendrogram and regional identity")

```

###Choosing a soft threshold 
In order to now perform WGCNA on our data, a "soft threshold" needs to be calculated. This produces some NAs (warning message suppressed)
```{r picking soft threshold, message=FALSE,warning=FALSE, cache=TRUE}
# Setting what soft thresholds (powers) should be tested for potential use as a soft threshold
powers <- c(c(1:10), seq(from = 12, to=20, by=2))

# Calculating soft thresholds for a signed network
sft <- pickSoftThreshold(datExpr, powerVector = powers,
                        verbose = 5,
                        networkType ="signed",
                        moreNetworkConcepts = F)

# Plotting the soft threshold diagnostics in order to choose an optimal soft threshold.
# Setting display margins
par(mfrow = c(1,1))
cex1 = 0.9

# Scale-free topology fit index as a function of the soft-thresholding power. We are looking to choose a soft threshold where the plot plateus out
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red")
abline(h=0.8,col="red") # this line corresponds to using an R^2 cut-off of h that could be used as a guide to picking a soft threshold

# Mean connectivity as a function of the soft-thresholding power. Again, we are looking at when the plot begins leveling out.
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")

# Writing out these plots
pdf(file = file.path(file_dir, "WGCNA.softpower.pdf"), width = 6, height = 6)
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red")
abline(h=0.8,col="red")

plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
dev.off()
```

### Performing topological overlap
A soft threshold of 10 was chosen based on the scale independence plot beginning to finally level out at that point for good. The dip from 8 to 9 prevented me from choosing 8. We choose a topology cut-off of 0.80, which is arbitary but similar to the number suggested in WGCNA tutorials (see "A General Framework for Weighted Gene CoExpression Network Analysis", 2005).
```{r performing topological overlap, cache=TRUE}
# Choosing a power of 10 based on Scale independence plot and 0.8 threshold
softPower <- 10
adjacency <- adjacency(datExpr, power = softPower, type = "signed")

# Turn adjacency into topological overlap
TOM <- TOMsimilarity(adjacency, TOMType = "signed")
dissTOM <- 1-TOM

# Call the hierarchical clustering function
geneTree <- flashClust(as.dist(dissTOM), method = "average")

# Plot the resulting clustering tree (dendrogram)
par(mfrow=c(1,1))
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.01)

# We like large modules, so we set the minimum module size relatively high:
minModuleSize <- 20

# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, 
                            deepSplit = T, pamRespectsDendro = FALSE, 
                            minClusterSize = minModuleSize)
table(dynamicMods)

# Convert numeric lables into colors
dynamicColors <- labels2colors(dynamicMods)
table(dynamicColors)

# And plot the gene dendrogram with module colors annotated underneath
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut", 
                    dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, 
                    guideHang = 0.04, main = "Gene dendrogram and module colors")

```

### Relationships between modules
Now that we have identified modules in our data, we want to explore the relationships between those modules a little more. Note that grey and turquoise modules were removed in these analyses because turquoise contained all genes that did not get slotted in to a specific module and grey had only 4 genes (below our set module size)
```{r relationships between modules, cache=TRUE}
# Calculate eigengenes for each module
MEList <- moduleEigengenes(datExpr, colors = dynamicColors)
MEs <- MEList$eigengenes

# Remove grey and turquoise for ahead of attempting to look at dissimilarity between eigengenes
MEs <- MEs[,c(1:6,8:16,18)] 

# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs)

# Cluster module eigengenes
METree <- flashClust(as.dist(MEDiss), method = "average")

# Plot the result
par(mfrow=c(1,1))
plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")
MEDissThres = 0.2
abline(h=MEDissThres, col = "red") # Plot the cut line into the dendrogram. Branches falling below this line will be merged. It does not look like any will get merged, but we run it anyways.

# Call an automatic merging function
merge <- mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)

# Extracting the merged module colors
moduleColors <- merge$colors

# Extracting eigengenes of the new merged modules:
mergedMEs <- merge$newMEs

# Constructing numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs <- mergedMEs
MEs <- MEs[,c(1:6,8:17)] #remove turquoise and grey again, since they for some reason are not gone

# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs)

# Cluster module eigengenes
METree <- flashClust(as.dist(MEDiss), method = "average")

# Plot the result
par(mfrow=c(1,1))
plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")

# write out the final plot to a pdf
pdf(file = file.path(file_dir,"Module_Eigengene_Tree.pdf"), height = 6, width = 8)
plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")
dev.off()
```

### Gene enrichment analyses
Now that we had explored the modules relationship to each other, we wanted to see if the genes within each module were enriched for gene ontology, KEGG, and Reactome terms. Not all genes map through 'bitr' so warnings were supressed
```{r gene enrichment analyses, message=F, warning=FALSE, cache=T}
# Making a list of module names
modNames <- substring(names(MEs), 3)

# Correlating each genes expression profile with the module eigengenes in order to create module gene sets
geneModuleMembership <- as.data.frame(cor(datExpr, MEs, use = "p"))
# "For each module, we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile." - WGCNA tutorial

# Iteratively creating a list of module genesets to test. These are in ensembl ids
moduleGeneSets<-lapply(modNames,function(module){
  column = match(module, modNames)
  moduleGenes = moduleColors==module
  rownames(geneModuleMembership[moduleGenes,])
})
names(moduleGeneSets)<-modNames

# Trimming the module gene sets so that the final two digits after the "." are removed
moduleGeneSets.trimmed<-lapply(moduleGeneSets,function(x){
  str_split_fixed(x,"\\.",2)[,1]
})

# Looking up the ENTREZ id for each gene
moduleGeneSets.Entrez<-lapply(moduleGeneSets.trimmed,function(x){
  bitr(x,fromType="ENSEMBL",toType="ENTREZID",OrgDb="org.Mm.eg.db")$ENTREZID
})

```

### Running geneset enrichment analyses
```{r, eval=T, message=F, warning=FALSE,fig.align='center',fig.height=10,fig.width=18,fig.asp=0.25}
# Looking at KEGG Geneset Enrichment
ck<-compareCluster(geneCluster=moduleGeneSets.Entrez,
                   fun="enrichKEGG",organism="mmu",pvalueCutoff = 0.1)
plot(ck,showCategory=10)+ggtitle("KEGG geneset enrichment by module") #width = 0.5

# Looking at molecular function geneset enrichment
cg.mf<-compareCluster(geneCluster=moduleGeneSets.Entrez,fun="enrichGO",
                      OrgDb = "org.Mm.eg.db",pvalueCutoff = 0.01,ont="MF", readable = T)
plot(cg.mf,showCategory=20)+ggtitle("GO MF geneset enrichment by module")

# Looking at biological process geneset enrichment 
cg.bp<-compareCluster(geneCluster=moduleGeneSets.Entrez,fun="enrichGO",
                      OrgDb = "org.Mm.eg.db", pvalueCutoff = 0.01,ont="BP", readable = T)
plot(cg.bp,showCategory=15)+ggtitle("GO BP geneset enrichment by module")

cg.cc<-compareCluster(geneCluster=moduleGeneSets.Entrez,fun="enrichGO",
                      OrgDb = "org.Mm.eg.db", pvalueCutoff = 0.01,ont="CC", readable = T)
plot(cg.cc,showCategory=15)+ggtitle("GO CC geneset enrichment by module")

# Looking at reactome enrichment
cr<-compareCluster(geneCluster=moduleGeneSets.Entrez,fun="enrichPathway",
                   organism="mouse",pvalueCutoff = 0.1, readable = T)
plot(cr,showCategory=10)+ggtitle("Reactome geneset enrichment by module")

# Printing out the cluster enrichment diagrams. These will be useful for annotating the 
pdf(file.path(file_dir,"P7-all-module_geneset_enrichments.pdf"),width=18,height=10, useDingbats = F)
plot(ck,showCategory=10)+ggtitle("KEGG geneset enrichment by module")
plot(cr,showCategory=10)+ggtitle("Reactome geneset enrichment by module")
plot(cg.bp,showCategory=10)+ggtitle("GO BP geneset enrichment by module")
plot(cg.mf,showCategory=10)+ggtitle("GO MF geneset enrichment by module")
plot(cg.cc,showCategory = 10) + ggtitle("GO CC geneset enrichment by module")
dev.off()

```

### Writing out GO tables
In order to output all of our GO data in a relatively simple way, all gene enrichment results were written out to a tables
```{r Getting outputs of cluster tables, eval=T,message=FALSE,warning=FALSE, cache=TRUE}

# Gene enrichment tables are found in 'ck','cg.mf','cg.bp','cg.cc', 'cr'
# summary() allows for the S4 objects created by clusterProfiler to be converted in to dataframes
kegg.table <- summary(ck)

# Had to make my own KEGG conversion function
kegg.table.clone <- kegg.table
for(i in 1:nrow(kegg.table.clone)){
  gene.str <- kegg.table.clone[i,9]
  gene.split <- unlist(strsplit(gene.str,split = "/"))
  genes.symbol <- bitr(gene.split, fromType = "ENTREZID", toType = "SYMBOL", OrgDb = 'org.Mm.eg.db')
  converted.gene.string <- paste0(genes.symbol$SYMBOL,collapse = "/")
  kegg.table.clone[i,9] <- converted.gene.string
}

# Easy enough to do for the others when you can set it at 'readable' in the functions above
go.molecular.function.table <- summary(cg.mf)
go.bio.process.table <- summary(cg.bp)
go.cell.comp.table <- summary(cg.cc)
reactome.table <- summary(cr)

# writing out tables
write.table(x = kegg.table.clone, file = file.path(file_dir,"WGCNA.KEGG.table.txt"), quote = F, sep = "\t", row.names = F)

write.table(x = go.molecular.function.table, file = file.path(file_dir,"GO.MF.table.txt"), quote = F, sep = "\t", row.names = F)

write.table(x = go.bio.process.table, file = file.path(file_dir,"GO.BP.table.txt"), quote = F, sep = "\t", row.names = F)

write.table(x = go.cell.comp.table, file = file.path(file_dir,"GO.CC.table.txt"), quote = F, sep = "\t", row.names = F)

write.table(x = reactome.table, file = file.path(file_dir,"Reactome.table.txt"), quote = F, sep = "\t", row.names = F)

```

### Writing out gene table 
In order to use them in downstream analyses, a table was written out in order to show all the genes in each module identified in our analysis (Table S6)
```{r writing a table out of genes in modules, eval=T,warning=F,message=FALSE}
library(reshape2)
# Turn the 'moduleGeneSets' list in to a data.frame, copy the columns and rename them 'module' and 'gene', select only the 'module' and 'gene' columns
module.df <- ldply(moduleGeneSets, data.frame) %>%
  mutate(module = .id, gene = X..i..) %>%
  dplyr::select(module, gene)

# Add id numbers in in order to unmelt the data and make sure the row lengths are the same for all columns
module.df$id = numeric(nrow(module.df))
for (i in 1:nrow(module.df)){
 dat_temp <- module.df[1:i,]
 module.df[i,]$id <- nrow(dat_temp[dat_temp$module == module.df[i,]$module,])
}

# Unmelt the data using "dcast" and then remove the 'id' column that helped you unmelt the data
module.genes.df <- dcast(module.df, id~module, value.var = "gene") %>%
  dplyr::select(-id)

# Replace all gene IDs with the mouse gene symbols
for(i in 1:ncol(module.genes.df)){
   geneIDs <- module.genes.df[,i][!is.na(module.genes.df[,i])]
   geneSymbols <- as.character(paste0("",lookupGeneName(eset = P7.dat.filter, geneIDs)))
   n <- 438 - length(geneSymbols)
   toAdd <- rep("",n)
   coltoAdd <- c(geneSymbols,toAdd)
   module.genes.df[,i] <- coltoAdd # adding a ' to gene names so they are not changed in Excel
}

# Write out the table
write.table(module.genes.df, file = file.path(file_dir,"WGCNA.module.gene.lists.txt"), quote = F, row.names = F, sep = "\t")
```

### Correlation of modules with traits 
Now that we have looked at enrichment for GO terms in our modules, we want to correlate the modules with cellular traitis, specifically subset cluster in order to find modules that differentiation between DA populations at P7
```{r Making a correlation table with all clusters called in P7 data}
# The first step to find these correlations is to make a trait table
# Copying over the datTraits dataframe
datTraits.new <- datTraits

datTraits.new$region <- as.factor(datTraits.new$region)
datTraits.new$subset.cluster <- as.factor(datTraits.new$subset.cluster)

# Using these for loops to create additional columns for each level of a trait, which will read "True" or "False" for each trait in each cell
for (l in levels(datTraits.new$subset.cluster)) {
  datTraits.new[[l]]<-datTraits.new$subset.cluster == l
}

for (l in levels(datTraits.new$region)){
  datTraits.new[[l]] <- datTraits.new$region == l
}

# Factoring the T/F columns in to a binary 0/1 classification system
datTraitsFactor<-datTraits.new
datTraitsFactor$region<-as.numeric(as.factor(datTraits.new$region))
datTraitsFactor$subset.cluster<-as.numeric(as.factor(datTraits.new$subset.cluster))

for (l in levels(datTraits.new$subset.cluster)){
  datTraitsFactor[[l]]<-as.numeric(datTraits.new[[l]])
}

for (l in levels(datTraits.new$region)){
  datTraitsFactor[[l]]<-as.numeric(datTraits.new[[l]])
}

# Keeping only the columns for subset clusters
datTraitsFactorCor <- datTraitsFactor[,c(6:14)]

# Correlating the module eigengenes with subset cluster identity
moduleTraitCor <- cor(MEs, datTraitsFactorCor)

# Getting a p-value for each correlation
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples)

# Plotting the relationship between learned gene co-expression modules and parameterizations
pdf(file = file.path(file_dir,"P7_module-trait_relationships-all.pdf"), width = 10, height = 8)
par(mfrow=c(1,1))
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
signif(moduleTraitPvalue, 1), ")", sep = "")
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3))

labeledHeatmap(Matrix = t(moduleTraitCor),
               xLabels = names(MEs), yLabels = names(datTraitsFactorCor),
               ySymbols = names(datTraitsFactorCor),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = t(textMatrix),
               setStdMargins = T, cex.text = 0.5, zlim = c(-1,1),
               main = paste("Module-trait relationships"))
dev.off()

par(mfrow=c(1,1))
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
signif(moduleTraitPvalue, 1), ")", sep = "")
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3))

labeledHeatmap(Matrix = t(moduleTraitCor),
               xLabels = names(MEs), yLabels = names(datTraitsFactorCor),
               ySymbols = names(datTraitsFactorCor),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = t(textMatrix),
               setStdMargins = T, cex.text = 0.5, zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```

### Correlate eigengenes with tSNE 
We then wanted to correlate the modules eigengenes with tSNE position and plot 7 that were found to be significantly correlated with a P7 subset cluster
```{r}
# Correlating tSNE position with eigenvectors
eigentSNE<-cbind(pData(P7.dat.WGCNA[,keepSamples])$tSNE1_pos,pData(P7.dat.WGCNA[,keepSamples,])$tSNE2_pos,MEs)
colnames(eigentSNE)<-c("tSNE1_pos","tSNE2_pos",modNames)

eigentSNE.melt<-melt(eigentSNE,id.vars=c("tSNE1_pos","tSNE2_pos"))

# Melting the data and selecting only the modules we want to plot 
eigentSNE.melt.subset <- eigentSNE.melt %>%
  filter(variable == "yellow" | variable == "green" | variable == "salmon" | variable == "brown" | variable == 
           "midnightblue" | variable == "blue" | variable == "lightcyan")
  
p2.subset<-ggplot(eigentSNE.melt.subset) + geom_point(aes(x=tSNE1_pos,y=tSNE2_pos,color=value),size=1) + monocle:::monocle_theme_opts() + scale_color_gradient2(low="blue",mid="white",high="red",limits=c(-0.1,0.1),oob=squish)

p2.subset.all <- p2.subset + facet_wrap("variable", nrow = 2) + ggtitle("WGCNA module eigengene expression across single P7 DA neurons") + coord_fixed(ratio = 1) + xlab("t-SNE 1") + ylab("t-SNE 2") + theme(plot.title = element_text(hjust = 0.5))

pdf(file = file.path(file_dir,"Eigen-tSNEplots-subset.pdf"), width = 10, height = 4)
p2.subset.all
dev.off()

p2.subset.all
```

### Save modules
Save modules that were enriched for PD genesets
```{r save park geneset}
saveRDS(moduleGeneSets$green, file = file.path(file_dir,"WGCNA.ParkGeneSet.Green.rds"))
saveRDS(moduleGeneSets$brown, file = file.path(file_dir,"WGCNA.ParkGeneSet.Brown.rds"))
```

###Session Info
```{r session infomation}
sessionInfo()
```
